---
title: "Data Cleaning Log"
author: "Joanna D."
date: "2025-10-22"
output: html_document
editor_options: 
  markdown: 
    wrap: 72
---

> The datasets used in this project are the first quarter bike-share
> trips from 2019 and 2020, and are sourced from [Cyclistic's (Divvy)
> public data bucket](https://divvybikes.com/system-data). The data was
> made available for public viewing and use, subject to conditions,
> under [this license](https://divvybikes.com/data-license-agreement).
> The 2019 Q1 dataset is referred to as Divvy_Trips_2019_Q1,
> Trips_2019_Q1, etc.; and the 2020 Q1 dataset is referred to as
> Divvy_Trips_2020_Q1, Trips_2020_Q1, etc. All data cleaning listed in
> this document was done in Google Sheets and RStudio unless otherwise
> stated.

# Google Sheets

## July 8, 2025

-   Made a copy of original 2019 Q1 dataset and named it `Trips_2019_Q1`
-   Auto-adjusted column width to fit contents
-   Froze row 1 and applied bold formatting (column headers)
-   Deleted extra empty rows (365071 and 365072) to accurately search
    for blanks
-   Created new sheet in 2019 Q1: `formulas`

## July 18, 2025

-   Checked for duplicates and white spaces to be trimmed; none found
-   Checked and verified that there were no `trip_id`s that were not
    equal to 8, as each ID must be exactly 8 characters long:
    `=(LEN(A2:A365070) <> 8) = TRUE`
-   Checked and verified that all `start_time`s and `end_time`s are
    either 18 or 19 characters with formula:
    `=(ISBETWEEN(LEN(B2:B365070), 18, 19)) = FALSE`
-   Checked and verified that all `bikeid`s are between 1 and 4
    characters with formula:
    `=(ISBETWEEN(LEN(D2:D365070), 1, 4)) = FALSE`
-   Used **Find** to check and verify that there are no decimals or
    percentages in the `bikeid` column.
-   Applied standardized date-time format to `start_time` and `end_time`
    columns: `2000-01-01 0:00:00`
-   Applied number format to `tripduration` column, then decreased
    decimal places to ensure all entries are *whole* numbers.
    -   Some were whole, some were decimals — the decimals only had .00
-   Applied number format to `from_station_id` column and decrease
    decimal places to ensure all entries are whole numbers.
-   In `formulas` sheet, began checking if `from_station_name` contained
    correct bike-share stations…
    -   Process
        1.  In the `formulas` sheet, I used the `UNIQUE` function to
            print all unique `from_station_name`s in column `A`, which I
            named `From Station`.
        2.  In column `B`, which I labeled `Exists?`, I auto-filled
            check-boxes in all cells so I could keep track of which
            stations I’ve checked already.
        3.  I used the Divvy System Map and Google Maps to confirm
            whether the station exists or not and is spelled correctly
            in the dataset.
        4.  Any rows with station names that needed to be revised were
            colored in cyan and had a comment on the station name that
            detailed what needed to be changed.
        5.  Any rows with stations that existed in the past but were no
            longer available OR that needed to be removed entirely
            (trips taken by employees servicing bikes) were colored in
            red and a comment on the station name that detailed what
            needed to be changed.
        6.  The row that I left off at was colored in yellow so I could
            easily pick up where I left off.

## July 21, 2025

-   In `formulas` sheet, continued checking if `from_station_name`
    contains correct bike-share stations…
    -   ❗Despite the station sign in person saying "Loomis St & Taylor
        St", the Loomis & Taylor station is actually off of Throop St &
        Taylor St and was changed to the latter name.

## July 22, 2025

-   In `formulas` sheet, continued checking if `from_station_name`
    contains correct bike-share stations…
    -   Copied and pasted `From Station`s as text to keep the station
        names and remove the `UNIQUE` formula.
    -   Sorted sheet by column `A` from A to Z.
    -   Added `To Station` column (`C`) because I forgot I need to check
        that simultaneously (🥹). Filled column with `UNIQUE`
        `to_station_name`s from the dataset.
    -   Added `Exists?` with check-boxes in column `D`.
    -   Sorted range of `To Station` names by column `C` from A to Z.

## July 24, 2025

-   In `formulas` sheet, continued checking if `from_station_name` and
    `to_station_name` contain correct bike-share stations…
    -   Station **Ashland Ave & 69th St** does not appear to exist on
        either G Maps or Divvy Map
        -   Accessed the dataset: [**Divvy Bicycle Stations -
            Historical**](https://data.cityofchicago.org/Transportation/Divvy-Bicycle-Stations-Historical/eq45-8inv/about_data)
            (from City of Chicago data) to confirm whether the station
            existed in the past.
            -   The dataset shows that the station existed until about
                October of 2020.
    -   Station **Ashland Ave & 21st St** does not appear to exist on
        either G Maps or Divvy Map
        -   Repeated steps for Ashland Ave & 69th St, which shows that
            the Ashland Ave & 21st St station existed until about late
            May of 2019.
    -   Station **Chicago Ave & Dempster St** does not appear to exist
        on Divvy Map, appears as “Permanently closed” on G Maps
        -   Repeated previous steps of checking City of Chicago
            historical Divvy station data, which shows that Chicago Ave
            & Dempster St existed as a station until about October of
            2020. 

## July 25, 2025

-   In `formulas` sheet, continued checking if `from_station_name` and
    `to_station_name` contain correct bike-share stations…
    -   Station **Clark St & Randolph St** does not appear to exist on
        Divvy Map or G Maps
        -   Repeated previous steps of checking City of Chicago
            historical Divvy station data, which shows that Clark St &
            Randolph St existed as a station until about September of
            2020. 
    -   Station **Cornell Ave & Hyde Park Blvd** does not appear to
        exist on Divvy Map, but it does on G Maps and Apple Map.
        -   G Maps and Apple Map street views show the station name as
            Cornell Ave & 50th Pl instead (May - June 2024).
        -   Divvy Map shows there are e-bikes there, but not a station.
        -   Repeated previous steps of checking City of Chicago
            historical Divvy station data, which shows that Cornell Ave
            & Hyde Park Blvd existed as a station until about September
            of 2020.
            -   [The City of Chicago website shows in 2014 that Divvy
                planned to build a station at Cornell Ave & Hyde Park
                Blvd.](https://www.chicago.gov/city/en/depts/cdot/provdrs/bike/news/2014/aug/DivvyExpansion.html)
    -   Station **Cornell Dr & Hayes Dr** does not appear to exist on
        Divvy Map, and is listed as “Permanently closed” on G Maps.
        -   G Maps street view shows it existed in June 2019 through
            July 2020, but was no longer there in August 2024.
    -   Station **Cottage Grove Ave & 63rd St** does not appear to exist
        on Divvy Map, but it does on G Maps and Apple Map.
        -   Repeated previous steps of checking City of Chicago
            historical Divvy station data, which shows that Cottage
            Grove Ave & 63rd St existed as a station up until about
            September of 2020
            -   G Maps and Apple Map street views conflict with this, as
                they both show that the station was still present and
                being used between May and June of 2024.
    -   Station **Damen Ave & Coulter St** does not appear to exist on
        Divvy Map, but it does on G Maps and Apple Map.
        -   Repeated previous steps of checking City of Chicago
            historical Divvy station data, which shows that Damen Ave &
            Coulter St existed as a station up until about September of

            2020. 

            -   G Maps and Apple Map street views conflict with this, as
                they both show that the station was still present and
                being used between November 2021 and June 2024.
    -   Station **DIVVY Map Frame B/C Station** is actually called the
        **Divvy Service Warehouse**
        -   Found this by searching for DIVVY Map Frame B/C Station in
            the Divvy Bicycle Stations - Historical dataset and
            retrieving the location’s latitude and longitude, then
            entering it into G Maps.
        -   To be removed.

## July 28, 2025

-   In `formulas` sheet, continued checking if `from_station_name` and
    `to_station_name` contain correct bike-share stations…
    -   Continuing from previous situation of Divvy Service Warehouse
        -   Station DIVVY CASSETTE REPAIR MOBILE STATION (41.895543
            -87.602921) is unmarked, but appears to be related to bike
            maintenance, therefore will remove.

## July 29, 2025

-   In `formulas` sheet, continued checking if `from_station_name` and
    `to_station_name` contain correct bike-share stations…

## July 30, 2025

-   In `formulas` sheet, continued checking if `from_station_name` and
    `to_station_name` contain correct bike-share stations…

## July 31, 2025

-   In `formulas` sheet, continued checking if `from_station_name` and
    `to_station_name` contain correct bike-share stations…
    -   July 31, 2025 10:21 AM **Station name checks complete.**
    -   Updating `from_station_name` and `to_station_name` in
        `Divvy_Trips_2019_Q1`...
        -   Replaced 83 instances of Archer (Damen) Ave & 37th St (\*)
            with **Archer (Damen) Ave & 37th St**
        -   Replaced 1,079 instances of Ashland Ave & Lake St (Temp)
            with **Ashland Ave & Lake St**
        -   Replaced 2,248 instances of Bissell St & Armitage Ave with
            **Bissell St & Armitage Ave**
        -   Replaced 598 instances of Buckingham Fountain (Temp) with
            **Buckingham Fountain**
        -   Replaced 1,589 instances of Burling St (Halsted) & Diversey
            Pkwy (Temp) with **Burling St (Halsted) & Diversey Pkwy**
        -   Replaced 4,017 instances of Canal St & Monroe St (\*) with
            **Canal St & Monroe St**
        -   Replaced 5,857 instances of Clinton St & Jackson Blvd (\*)
            with **Clinton St & Jackson Blvd**
        -   Replaced 555 instances of Clinton St & Polk St (\*) with
            **Clinton St & Polk St**
        -   Replaced 259 instances of Damen Ave & Walnut (Lake) St (\*)
            with **Damen Ave & Lake St**
        -   Replaced 1,524 instances of Dearborn St & Van Buren St (\*)
            with **Dearborn St & Van Buren St**
        -   Replaced 911 instances of Dusable Harbor with **DuSable
            Harbor**
        -   Replaced 435 instances of Eastlake Ter & Rogers Ave with
            **Eastlake Ter & Howard St**
        -   Replaced 20 instances of Eggleston Ave & 69th St (\*) with
            **Eggleston Ave & 69th St**
        -   Replaced 1,965 instances of Fairbanks St & Superior St (\*)
            with **Fairbanks St & Superior St**
        -   Replaced 2,373 instances of Financial Pl & Ida B Wells Dr
            (Temp) with **Financial Pl & Ida B Wells Dr**
        -   Replaced 2,772 instances of Green St & Randolph St with
            \*\*Green St & Randolph St (\*)\*\*
        -   Replaced 140 instances of Halsted St & 35th St (\*) with
            **Halsted St & 35th St**
        -   Replaced 2,666 instances of Halsted St & Clybourn Ave (\*)
            with **Halsted St & Clybourn Ave**
        -   Replaced 1,687 instances of LaSalle Dr & Huron St (\*) with
            **LaSalle Dr & Huron St**
        -   Replaced 208 instances of Leavitt St & Belmont Ave (\*) with
            **Leavitt St & Belmont Ave**
        -   Replaced 383 instances of Leavitt St & Division St (\*) with
            **Leavitt St & Division St**
        -   Replaced 1,659 instances of Lincoln Ave & Roscoe St with
            \*\*Lincoln Ave & Roscoe St (\*)\*\*
        -   Replaced 1,256 instances of Lincoln Ave & Belmont Ave with
            **Lincoln Ave & Belmont Ave (N/A)**
        -   Replaced 915 instances of Loomis St & Taylor St with
            **Throop St & Taylor St**
        -   Replaced 42 instances of MLK Jr Dr & 56th St (\*) with **MLK
            Jr Dr & 56th St**
        -   Replaced 4,636 instances of Morgan St & Lake St with
            \*\*Morgan St & Lake St (\*)\*\*
        -   Replaced 284 instances of Museum of Science and Industry
            with **Griffin Museum of Science and Industry**
        -   Replaced 1,066 instances of Orleans St & Elm St (\*) with
            **Orleans St & Elm St**
        -   Replaced 2,137 instances of Orleans St & Hubbard St (\*)
            with **Orleans St & Hubbard St**
        -   Replaced 87 instances of Pulaski Rd & Eddy St (Temp) with
            **Pulaski Rd & Eddy St**
        -   Replaced 982 instances of Racine Ave & Washington Blvd (\*)
            with **Racine Ave & Washington Blvd**
        -   Replaced 2,102 instances of Sangamon St & Washington Blvd
            (\*) with **Sangamon St & Washington Blvd**
        -   Replaced 121 instances of Smith Park (\*) with **Smith
            Park**
        -   Replaced 7 instances of Stewart Ave & 63rd St (\*) with
            **Stewart Ave & 63rd St**
        -   Replaced 4,729 instances of Streeter Dr & Grand Ave with
            **Navy Pier**
        -   Replaced 245 instances of Washtenaw Ave & Ogden Ave (\*)
            with **Washtenaw Ave & Ogden Ave**
        -   Replaced 568 instances of Wentworth Ave & Cermak Rd (Temp)
            with \*\*Wentworth Ave & Cermak Rd (\*)\*\*
        -   Replaced 150 instances of Western Ave & Fillmore St (\*)
            with **Western Ave & Fillmore St**
        -   Repalced 640 instances of Wolcott (Ravenswood) Ave &
            Montrose Ave (\*) with **Wolcott (Ravenswood) Ave & Montrose
            Ave**
        -   Replaced 810 instances of Wood St & Chicago Ave (\*) with
            **Wood St & Chicago Ave**
        -   Replaced 1,576 instances of Wood St & Taylor St with **Wood
            St & Taylor St (Temp)**
        -   Replaced 300 instances of Ashland Ave & 21st St with
            **Ashland Ave & 21st St (N/A)**
        -   Replaced 23 instances of Ashland Ave & 69th St with
            **Ashland Ave & 69th St (N/A)**
        -   Replaced 408 instances of Chicago Ave & Dempster St with
            **Chicago Ave & Dempster St (N/A)**
        -   Replaced 4,210 instances of Clark St & Randolph St with
            **Clark St & Randolph St (N/A)**
        -   Re-checked station Cornell Ave & Hyde Park Blvd, which [did
            not appear as a station](##july-25-2025) until some time
            between July 25, 2025 and July 31, 2025. No need for name
            edits or removal. (close call…)
        -   Replaced 7 instances of Cornell Dr & Hayes Dr with **Cornell
            Dr & Hayes Dr (N/A)**
        -   Re-checked station Cottage Grove Ave & 63rd St, which [did
            not appear as a station](##july-25-2025) until some time
            between July 25, 2025 and July 31, 2025. No need for name
            edits or removal. (another one!)
        -   Replaced 38 instances of Damen Ave & Coulter St with **Damen
            Ave & Coulter St (N/A)**
        -   Removed all trips to and from the following stations: -
            DIVVY Map Frame B/C Station - DIVVY CASSETTE REPAIR MOBILE
            STATION - `trip_id`s of deleted rows - 21758019 - 21764967 -
            21842257 - 21902493 - 21915949 - 21920842 - 21926356 -
            21949128 - 21967146 - 21979478 - 21981589 - 21990694 -
            21994800 - 21999735 - 22002795 - 22003744 - 22017886 -
            22020622 - 22027378 - 22067197 - 22067200 - 22095022 -
            22095154 - 22100544 - 22108999 - 22109359 - 22165514
        -   Replaced 351 instances of Drake Ave & Fullerton Ave with
            **Drake Ave & Fullerton Ave (N/A)**
        -   Replaced 2 instances of Elizabeth St & 59th St with
            **Elizabeth St & 59th St (N/A)**
        -   Replaced 3,818 instances of Franklin St & Quincy St with
            **Franklin St & Quincy St (N/A)**
        -   **Left off at Lake Shore Dr & Ohio St**

## August 1, 2025

-   Fixing `from_station_name` and `to_station_name` in
    `Divvy_Trips_2019_Q1`... - Replaced 1,433 instances of Lake Shore Dr
    & Ohio St with **Lake Shore Dr & Ohio St (N/A)** - Replaced 1,256
    instances of Lincoln Ave & Belmont Ave with **Lincoln Ave & Belmont
    Ave (N/A)** - Replaced 3,368 instances of Marshfield Ave & Cortland
    St with **Marshfield Ave & Cortland St (N/A)** - Replaced 3,455
    instances of McClurg Ct & Illinois St with **McClurg Ct & Illinois
    St (N/A)** - Replaced 1,896 instances of Milwaukee Ave & Wabansia
    Ave with **Milwaukee Ave & Wabansia Ave (N/A)** - Re-checked station
    MLK Jr Dr & Pershing Rd, which did not appear as a station until
    some time between July 29, 2025 and August 1, 2025. No need for name
    edits or removal. (!!!) - For checking missing stations, I used the
    Divvy Bicycle Stations dataset (a list of current stations) to
    confirm that the stations I marked in red in the `formulas` sheet
    are no longer present. - Replaced 1,731 instances of Racine Ave (May
    St) & Fulton St with **Racine Ave (May St) & Fulton St (N/A)**
    -   August 1, 2025 10:53 AM **Updating station names complete**
-   In `Divvy_Trips_2019_Q1`, renamed columns:
    -   D: `bikeid` → `bike_id`
    -   E: `tripduration` → `trip_duration`
    -   J: `usertype` → `user_type`
    -   L: `birthyear` → `birth_year`
-   Auto-adjusted column length to fit content
-   Updated column E: `trip_duration` → `trip_duration_secs`
-   Inserted new column to right of column E: `trip_duration_h_mm_ss`
    -   In column F, converted all trip durations from seconds to hours,
        minutes, and seconds in this format: 0:00:00
        -   Used the formula `=[trip_duration_secs]/(24*60*60)` and
            formatted the output as a decimal with 6 places, and then
            converted the decimal to the duration format.
-   Hid column E `trip_duration_secs`
-   In column L `gender`, used filters to only show blanks.
    -   Filled blanks with “N/A”
-   In column M `birth_year`, used filters to only show blanks.
    -   Filled blanks with “N/A”
-   Noticed that the earliest birth year in the dataset is 1900, which
    as of 2025 was **about 125 years ago.** 🥸
    -   There are 69 records where the user entered the birth year 1900.
    -   Noticed and corrected a station name error: Lincoln Ave &
        Belmont Ave (N/A) (N/A) → **Lincoln Ave & Belmont Ave (N/A)**
    -   Birth years of concern:
        -   1900 (125 yo, 69 records, all subscribers)
        -   1901 (124 yo, 2 records, both customers)
        -   1918 (107 yo, 40 records, all subscribers)
        -   1921 (104 yo, 60 records, all subscribers)
    -   **Leaving off note: consider possibly filling in abnormal birth
        years with “N/A”?**

## August 4, 2025

-   Continued from the previous situation of abnormal birth years…
    -   Replaced abnormal birth years with N/A
-   Applied bold formatting row 1: column headers
-   August 4, 2025 10:07 AM **Switching over to cleaning**
    `Trips_2020_Q1` dataset
    -   Froze and applied bold formatting to row 1: column headers
    -   Created sheet `formulas`
    -   Auto-adjusted column width
    -   Adjusted column order
        -   From:
            -   **ride_id → rideable_type → started_at → ended_at →
                start_station_name → start_station_id → end_station_name
                → end_station_id → start_lat → start_lng → end_lat →
                end_lng → member_casual**
        -   To:
            -   **ride_id → rideable_type → started_at → ended_at →
                start_station_id → start_station_name → end_station_id →
                end_station_name → start_lat → start_lng → end_lat →
                end_lng → member_casual**
    -   Sorted sheet by column C (`started_at`) from A to Z

## August 11, 2025

-   Standardized column names across data sets
    -   In `Divvy_Trips_2020_Q1`:
        -   `A1`: `ride_id` → `trip_id`
        -   `C1`: `started_at` → `start_time`
        -   `D1`: `ended_at` → `end_time`
        -   `M1`: `member_casual` → `user_type`
    -   In `Divvy_Trips_2019_Q1`:
        -   `G1`: `from_station_id` → `start_station_id`
        -   `H1`: `from_station_name` → `start_station_name`
        -   `I1`: `to_station_id` → `end_station_id`
        -   `J1`: `to_station_id` → `end_station_name`
-   In the 2019 Q1 dataset, renamed `user_type`s:
    -   Subscriber → Member
    -   Customer → Casual
-   In the 2020 Q1 dataset
    -   Fixed `user_type` names:

        -   casual → Casual
        -   member → Member

    -   Inserted one column to the right of column D `end_time`: column
        E, `trip_duration_h_mm_ss`

    -   In column E, inserted formula for each cell to calculate trip
        duration (`end_time - start_time`), formatted as a duration in
        hours, minutes, and seconds (0:00:000)

    -   Applied filters and checked column E, noticed there were trip
        durations less than 60 seconds (0:01:00), some even in the
        negatives…!

        -   [The data set's source stated that trips taken by staff and
            trips that are below 60 seconds in length were removed from
            the data sets they
            share](https://divvybikes.com/system-data), so any that meet
            this criteria must be removed if still present (they still
            are in the 2020 Q1 dataset as of August 11, 2025 8:11 AM
            lol).

    -   Deleted blank rows 426889 and 426890

    -   Applied filters to only show negative trip durations (**Text
        contains: -**); all 116 trips had a start station of “HQ QR”,
        and 115 of them had an end station of “HQ QR” — the other one
        end station was blank.

        -   Looked up the latitude and longitude of the HQ QR station —
            it’s the **Divvy Service Warehouse**.
        -   Deleted all trips with negative durations because they were
            all taken by staff servicing bikes at/around the warehouse
            AND less than 60 seconds.

    -   Applied filters to only show trip durations less than 60 seconds
        (Text starts with: 0:00)

        -   Started deleting all trips less than 60 seconds…technical
            difficulties 🥔🔨

## August 18, 2025

-   Still in 2020 Q1 Trips: Finished deleting all trips less than 60
    seconds! 🎊 August 18, 2025 9:35 AM
-   Applied filters to only show trips with start stations of “HQ QR” —
    4 displayed, end station is the same
    -   Trips deleted August 18, 2025 9:40 AM
-   Applied filters to only show trips with end stations of “HQ QR” — 1
    displayed, start station is Adler Planetarium
    -   The trip duration is 62 hrs, 23 mins, and 34 secs, which is
        about 2.5 days.
    -   Trip deleted
-   Trip with an abnormal duration of **2,607 hours, 30 minutes, and 24
    seconds, which is about 108 days!!!**
    -   **WILL SWING BACK TO THIS**
        -   <https://help.divvybikes.com/hc/en-us/articles/360033484791-What-if-I-keep-a-bike-out-too-long>
-   Inserted new column C: `start_day`
    -   In column C, filled cells with formula to print the day of the
        week each trip started: `=TEXT(X#, “dddd”)`
    -   In 2019 Q1 Trips dataset, did the same (inserted new column B of
        the same name with the similar formulas in each cell below).
-   Back in the 2020 Q1 Trips dataset, in the `formulas` sheet I printed
    all unique start station names and end station names to repeat the
    process of checking the existence of stations.
    -   NTS: Use the 2019 Q1 Trips station names to help you quickly
        identify which still exist!!!!!!

## August 19, 2025

-   Trips_2020_Q1
    -   In `formulas` sheet, combined start and end station names in one
        column and used the `UNIQUE` function to get all unique start
        and end station names so I can check them simultaneously. Wish I
        did that the first time! The more you know 🌠
    -   **Checking validity of stations**
        -   Left off at row 33 — Austin Blvd & Lake St

## August 20, 2025

-   Trips_2020_Q1
    -   **Checking validity of stations**; `formulas`
        -   To rename:
            -   Bissell St & Armitage Ave → \*\*Bissell St & Armitage
                Ave (\*)\*\*
            -   Burling St (Halsted) & Diversey Pkwy (Temp) → **Burling
                St & Diversey Pkwy**
            -   California Ave & Francis Pl → **California Ave & Francis
                Pl**
            -   Chicago Ave & Dempster St → **Chicago Ave & Dempster St
                (N/A)**
            -   Clark St & Randolph St → **Clark St & Randolph St
                (N/A)**
            -   Cornell Dr & Hayes Dr → **Cornell Dr & Hayes Dr (N/A)**
            -   Damen Ave & Coulter St → **Damen Ave & Coulter St
                (N/A)**
            -   Damen Ave & Walnut (Lake) St (\*) → **Damen Ave & Lake
                St**
            -   Drake Ave & Fullerton Ave → **Drake Ave & Fullerton Ave
                (N/A)**
            -   Dusable Harbor → **DuSable Harbor**
            -   Eastlake Ter & Rogers Ave → **Eastlake Ter & Howard St**
            -   Eggleston Ave & 69th St (\*) → **Eggleston Ave & 69th
                St**
            -   Elizabeth (May) St & Fulton St → **Elizabeth St & Fulton
                St**
            -   Elizabeth St & 59th St → **Elizabeth St & 59th St
                (N/A)**
            -   Franklin St & Adams St (Temp) → **Franklin St & Adams
                St**
            -   Green St & Randolph St → \*\*Green St & Randolph St
                (\*)\*\*
        -   To remove:
            -   HUBBARD ST BIKE CHECKING (LBS-WH-TEST)
        -   Left off at row 273: Humboldt Blvd & Armitage Ave

## August 26, 2025

-   Trips_2020_Q1
    -   **Checking validity of stations**; `formulas`
        -   To rename:
            -   Lake Shore Dr & Ohio St → **Lake Shore Dr & Ohio St
                (N/A)**
            -   Leavitt St & Belmont Ave (\*) → **Leavitt St & Belmont
                Ave**
            -   Leavitt St & Division St (\*) → **Leavitt St & Division
                St**
            -   Lincoln Ave & Belmont Ave → **Lincoln Ave & Belmont Ave
                (N/A)**
            -   Lincoln Ave & Roscoe St → \*\*Lincoln Ave & Roscoe St
                (\*)\*\*
            -   Marshfield Ave & Cortland St → **Marshfield Ave &
                Cortland St (N/A)**
            -   McClurg Ct & Illinois St → **McClurg Ct & Illinois St
                (N/A)**
            -   Milwaukee Ave & Wabansia Ave → **Milwaukee Ave &
                Wabansia Ave (N/A)**
            -   Morgan St & Lake St → \*\*Morgan St & Lake St (\*)\*\*
            -   Museum of Science and Industry → **Griffin Museum of
                Science and Industry**
            -   Pulaski Rd & Eddy St (Temp) → **Pulaski Rd & Eddy St**
            -   Racine Ave & Washington Blvd (\*) → **Racine Ave &
                Washington Blvd**
            -   Smith Park (\*) → **Smith Park**
        -   To remove:
            -   MTL-ECO5.1-01
                -   In this dataset, records with this station had both
                    start and end coordinates 41.9171,-87.7482, which
                    pointed to **4854 West Armitage Avenue, Chicago, IL
                    60639**. Oddly, when I searched for MTL-ECO5.1-01 in
                    the Divvy Bicycle Stations - Historical dataset,
                    rows with the coordinates 41.862949,-87.608585
                    returned, which pointed to **Huntington Bank
                    Pavilion at Northerly Island, 1300 South Linn White
                    Drive, Chicago, IL 60605**.
                -   Either way, this thing's gotta go. :]
        -   Left off at row 494: South Shore Dr & 67th St

## August 27, 2025

-   Trips_2020_Q1
    -   **Checking validity of stations**; `formulas`
        -   To rename:
            -   Stewart Ave & 63rd St (\*) → **Stewart Ave & 63rd St**
            -   Streeter Dr & Grand Ave → **Navy Pier**
            -   Throop (Loomis) St & Taylor St → **Throop St & Taylor
                St**
            -   Wentworth Ave & 24th St (Temp) → **Wentworth Ave & 24th
                St**
            -   Wentworth Ave & Cermak Rd (Temp) → \*\*Wentworth Ave &
                Cermak Rd (\*)\*\*
            -   Western Ave & Fillmore St (\*) → **Western Ave &
                Fillmore St**
            -   Wood St & Chicago Ave (\*) → **Wood St & Chicago Ave**
            -   Wood St & Taylor St → **Wood St & Taylor St (Temp)**
            -   Ashland Ave & 69th St → **Ashland Ave & 69th St (N/A)**
        -   August 27, 2025 12:37 PM Checking complete!
    -   Updating station names in `Divvy_Trips_2020_Q1`
        -   Replaced 3,263 instances of Bissell St & Armitage Ave with
            \*\*Bissell St & Armitage Ave (\*)\*\*
        -   Replaced 2,512 instances of Burling St (Halsted) & Diversey
            Pkwy (Temp) with **Burling St & Diversey Pkwy**
        -   Replaced 557 instances of California Ave & Francis Pl (Temp)
            with **California Ave & Francis Pl**

## September 2, 2025

-   Updating station names in `Divvy_Trips_2020_Q1`
    -   Replaced 493 instances of Chicago Ave & Dempster St with
        **Chicago Ave & Dempster St (N/A)**
    -   Replaced 4,736 instances of Clark St & Randolph St with **Clark
        St & Randolph St (N/A)**
    -   Replaced 17 instances of Cornell Dr & Hayes Dr with **Cornell Dr
        & Hayes Dr (N/A)**
    -   Replaced 86 instances of Damen Ave & Coulter St with **Damen Ave
        & Coulter St (N/A)**
    -   Replaced 427 instances of Damen Ave & Walnut (Lake) St (\*) with
        **Damen Ave & Lake St**
    -   Replaced 538 instances of Drake Ave & Fullerton Ave with **Drake
        Ave & Fullerton Ave (N/A)**
    -   Replaced 1,522 instances of Dusable Harbor with **DuSable
        Harbor**
    -   Replaced 722 instances of Eastlake Ter & Rogers Ave with
        **Eastlake Ter & Howard St**
    -   Replaced 2 instances of Eggleston Ave & 69th St (\*) with
        **Eggleston Ave & 69th St**
    -   Replaced 5 instances of Elizabeth St & 59th St with **Elizabeth
        St & 59th St (N/A)**
    -   Replaced 1,455 instances of Elizabeth (May) St & Fulton St with
        **Elizabeth St & Fulton St**
    -   Replaced 2,859 instances of Franklin St & Adams St with
        **Franklin St & Adams St**
    -   Replaced 2,813 instances of Green St & Randolph St with
        \*\*Green St & Randolph St (\*)\*\*
    -   Removed records with start and end stations **HUBBARD ST BIKE
        CHECKING (LBS-WH-TEST)**
    -   Replaced 2,352 instances of Lake Shore Dr & Ohio St with **Lake
        Shore Dr & Ohio St (N/A)**
    -   Replaced 495 instances of Leavitt St & Belmont Ave (\*) with
        **Leavitt St & Belmont Ave**
    -   Replaced 675 instances of Leavitt St & Division St (\*) with
        **Leavitt St & Division St**
    -   Replaced 1,497 instances of Lincoln Ave & Belmont Ave with
        **Lincoln Ave & Belmont Ave (N/A)**
    -   Replaced 1,919 instances of Lincoln Ave & Roscoe St with
        \*\*Lincoln Ave & Roscoe St (\*)\*\*
    -   Replaced 3,735 instances of Marshfield Ave & Cortland St with
        **Marshfield Ave & Cortland St (N/A)**
    -   Replaced 3,982 instances of McClurg Ct & Illinois St with
        **McClurg Ct & Illinois St (N/A)**
    -   Replaced 2,205 instances of Milwaukee Ave & Wabansia Ave with
        **Milwaukee Ave & Wabansia Ave (N/A)**
    -   Replaced 5,571 instances of Morgan St & Lake St with \*\*Morgan
        St & Lake St (\*)\*\*
    -   Removed records with start and end stations **MTL-ECO5.1-01**
    -   Replaced 396 instances of Museum of Science and Industry with
        **Griffin Museum of Science and Industry**
    -   Replaced 222 instances of Pulaski Rd & Eddy St (Temp) with
        **Pulaski Rd & Eddy St**
    -   Replaced 1,114 instances of Racine Ave & Washington Blvd (\*)
        with **Racine Ave & Washington Blvd**
    -   Replaced 184 instances of Smith Park (\*) with **Smith Park**
    -   Replaced 7 instances of Stewart Ave & 63rd St (\*) with
        **Stewart Ave & 63rd St**
    -   Replaced 5,180 instances of Streeter Dr & Grand Ave with **Navy
        Pier**
    -   Replaced 1,331 instances of Throop (Loomis) St & Taylor St with
        **Throop St & Taylor St**
    -   Replaced 287 instances of Wentworth Ave & 24th St (Temp) with
        **Wentworth Ave & 24th St**
    -   Replaced 751 instances of Wentworth Ave & Cermak Rd (Temp) with
        \*\*Wentworth Ave & Cermak Rd (\*)\*\*
    -   Replaced 518 instances of Western Ave & Fillmore St (\*) with
        **Western Ave & Fillmore St**
    -   Replaced 1,151 instances of Wood St & Chicago Ave (\*) with
        **Wood St & Chicago Ave**
    -   Replaced 1,637 instances of Wood St & Taylor St with **Wood St &
        Taylor St (Temp)**
    -   Replaced 3 instances of Ashland Ave & 69th St with **Ashland Ave
        & 69th St (N/A)**
    -   Done with station names September 2, 2025 11:33 AM

## September 3, 2025

-   Hid (**DID NOT DELETE**) columns that were not present in the other
    dataset September 3, 2025 11:57 AM
    -   `Trips_2019_Q1` — these columns were not present in
        `Trips_2020_Q1` and could not be obtained nor inferred.
        -   `birth_year`
        -   `gender`
    -   `Trips_2020_Q1` — these columns were not present in
        `Trips_2019_Q1` , and although they could have been obtained,
        they were not relevant for my analysis.
        -   `start_lat`
        -   `start_lng`
        -   `end_lat`
        -   `end_lng`
        -   `rideable_type`
            -   All entries in this column were “docked_bike”

## September 4, 2025

-   Swung back to `Trips_2019_Q1` to fix some formulas, column names,
    and formats.
    -   `bike_id` → `rideable_type`
        -   Converted all cells under the `rideable_type` column to
            plain text for easier merging in R. ✨
    -   Dropped `trip_duration_secs` column
    -   Cleared cells in `trip_duration_h_mm_ss` column
        -   `trip_duration_h_mm_ss` → `trip_duration`
            -   Did the same in `Trips_2020_Q1` .
        -   Changed duration formatting: 1:02:03 → 1 hr 2 min 3 sec
            -   Did the same in `Trips_2020_Q1` .
        -   Inserted formula in `trip_duration` column to calculate
            duration: `D#-C#`
-   Done; exported each dataset to .csv files. September 4, 2025 1:23 PM

# RStudio

## September 3 to \~9, 2025

> For more details on my analysis done in R, check out this [R
> script](https://github.com/jody-3/cyclistic-case-study/blob/main/Analysis_from_2025-09-03_to_2025-09-10.R).

-   Installed and loaded necessary packages:

```{r}
install.packages("tidyverse")
library(tidyverse)
library(conflicted)
```

-   Imported data set .csv files:

```{r}
# Changed working directory first with command: setwd("path/to/directory")
q1_2019 <- read_csv("Trips_2019_Q1.csv")
q1_2020 <- read_csv("Trips_2020_Q1.csv")
```

-   Converted `trip_id` and `rideable_type` columns to character types
    for proper stacking:

```{r}
q1_2019 <- mutate(q1_2019, trip_id = as.character(trip_id),
                  rideable_type = as.character(rideable_type))
```

-   Merged data frames:

```{r}
all_trips <- bind_rows(q1_2019, q1_2020)
```

-   Dropped `lat`, `long`, `birth_year`, and `gender` fields as this
    data was dropped beginning in 2020:

```{r}
all_trips <- all_trips %>%
  select(-c(start_lat, start_lng, end_lat, end_lng, birth_year, gender))
```

-   Added date, month, and year columns to allow for aggregation of ride
    data by measurements of time instead of at individual ride levels:

```{r}
all_trips$date <- as.Date(all_trips$start_time)
all_trips$month <- format(as.Date(all_trips$date), "%m")
all_trips$day_of_month <- format(as.Date(all_trips$date), "%d")
all_trips$year <- format(as.Date(all_trips$date), "%Y")
(all_trips <- rename(all_trips
                     ,day_of_week = start_day))
```

-   Converted `trip_duration` to numeric type, with duration in seconds
    only, for analysis:

```{r}
# Extracted hours, minutes, and seconds and stored as separate fields
all_trips <- all_trips %>% 
  mutate(
    hours = as.numeric(str_extract(trip_duration, "\\d+(?= hr)")),
    minutes = as.numeric(str_extract(trip_duration, "\\d+(?= min)")),
    seconds = as.numeric(str_extract(trip_duration, "\\d+(?= sec)"))
  )

# Converted hours, minutes, and seconds to seconds only and stored in a new field
all_trips <- all_trips %>% 
  mutate(
    duration_secs = hours * 3600 + minutes * 60 + seconds
  )

# Removed hours, minutes, and seconds columns as they're no longer needed
all_trips <- all_trips %>%
  select(-hours, -minutes, -seconds)
```
